---
title: "Analysis 1"
author: "Charlie Roth"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Packages and Data

I load the packages 'tidyverse' and 'GGally'. I then load my two datasets using read_csv() and name it "deer_pop" and "pop".

```{r}
library(tidyverse)
library(GGally)

deer_pop <- read_csv("deer_pop_data.csv")
pop <- read_csv("final_pop_data.csv")
```

Next, I use the function spread() on my "pop" dataset to get wide data instead of long data.

```{r}
data_wide <- spread(pop, county, total_pop)
```

I do the same for "deer_pop" and rename the "Years" column to "years".

```{r}
data_wide2 <- spread(deer_pop, ParkName, DeerNum)
data_wide2 <- data_wide2 |>
  rename(years = "Years")
```

I did this so I can combine the datasets by the "years" column.

```{r}
merged_data <- left_join(data_wide2, data_wide, by = "years")
```

Next, I rename the columns to shorter names for easier analysis.

```{r}
merged_data <- merged_data |>
  rename(AP_deer = "Arkansas Post National Memorial",
         PR_deer = "Pea Ridge National Military Park",
         WC_deer = "Wilson's Creek National Battlefield",
         AP_human = "Arkansas County, Arkansas",
         PR_human = "Benton County, Arkansas",
         WC_human = "Christian/Greene Counties, Missouri")
```

I use ggpairs to get a pairwise plot to see any correlation between the variables.

```{r}
ggpairs(merged_data, columns = 2:7, progress = F)
```

You can see that the human populations are all correlated which makes sense. Arkansas Post's deer population and human population surrounding it are not very correlated at a -0.397. Pea Ridge's populations are slightly correlated at 0.633 and Wilson's Creek populations are very correlated at 0.958.

To look further into this I decided to do linear regression models for each of the three parks.

```{r}
lm_model <- lm(AP_deer ~ AP_human, data = merged_data)
summary(lm_model)
```

From this model you can see that the Adjusted R-squared if negative and the p-value is high. This is most likely due to the small sample size.

```{r}
lm_model <- lm(PR_deer ~ PR_human, data = merged_data)
summary(lm_model)
```

Like Arkansas Post, Pea Ridge has a negative R-squared and a high p-value. Again I think this is most likely due to a small sample size.

```{r}
lm_model <- lm(WC_deer ~ WC_human, data = merged_data)
summary(lm_model)
```

Unlike the other two, Wilson's Creek has a high R-squared. Like the other two the p-value is not significant.

Based on the results from my analysis, I believe that the sample size is too small to make any conclusions on if deer population size is affected by human population size.