---
title: "Cleaning 1"
author: "Charlie Roth"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Packages and Data

I load the packages 'tidyverse' and 'viridisLite'. I then set my working directory with setwd() and loaded my data using read_csv() and named it "deer".

```{r}
library(tidyverse)

deer <- read_csv("HTLN_WhiteTailedDeer_Monitoring.csv")
```

## Variable Selection

I use select to remove the variables I won't be using for my project and create a new dataset, "deer2", with only the variables I will use.

```{r}
deer2 <- deer |>
  select(ParkName, BeginningTemperatureInCelsius, BeginningPrecipitation, Year, DeerNumber, VegetationType, GPSDate, GPSTime, LongitudeInDecimalDegrees, LatitudeInDecimalDegrees)
```

## Checking for NAs

I first use the function anyNA() to check for NAs in my dataset as a whole. The output result is "TRUE". This means that there are NAs in my dataset which I have to remove or fix. I then use the function anyNA() on each of my variables one by one until I get a "TRUE".

```{r}
anyNA(deer2)
anyNA(deer2$ParkName)
anyNA(deer2$BeginningTemperatureInCelsius)
anyNA(deer2$BeginningPrecipitation)
anyNA(deer2$Year)
anyNA(deer2$DeerNumber)
anyNA(deer2$VegetationType)
anyNA(deer2$GPSDate)
anyNA(deer2$GPSTime)
```

The result for the variable GPSTime was true so I remove the NAs using filter(). I then check for NAs in the whole dataset again to see if There are any more NAs in other variables.

```{r}
deer2 <- deer2 |>
  filter(!is.na(GPSTime))
anyNA(deer2)
```

The output is "FALSE" which means my dataset has no more NAs and I can stop the process of checking for them.

## Formatting

I decided to create a separate column of just month numbers using format() and as.Date().

```{r}
deer2$Month <- format(as.Date(deer2$GPSDate, format = "%m/%d/%Y"), "%m")
print(deer2)
```

I also renamed some of the columns that had very long names as you can see below.

```{r}
deer2 <- deer2 |>
  rename(BeginTempCelsius = BeginningTemperatureInCelsius,
         BeginPrecip = BeginningPrecipitation,
         DeerNum = DeerNumber,
         VegType = VegetationType,
         Long = LongitudeInDecimalDegrees,
         Lat = LatitudeInDecimalDegrees)
```

After cleaning and formatting my data, I create a .csv file with the new data that I will use in my EDA.

```{r}
df <- data.frame(deer2)

#write_csv(df, "C:/Users/asoph/OneDrive/Desktop/Spring 2025/DATA 205/Project/Cleaning/Cleaning 1/clean_deer.csv")
```
